# Digit Recognizer
This is a simple model to recognize grayscale images of digits 0-9
```{r}
# Loading required packages
library(caret)
library(e1071)

# Reading in data and viewing its summary
train=read.csv("../input/train.csv")
test=read.csv("../input/test.csv")
dim(train);dim(test)
summary(train)
```
The above data is loaded from the kaggle competition page https://www.kaggle.com/c/digit-recognizer/data

```{r}
test$label=1                                # Adding label column in the test dataset
# Combining data
comb_data=rbind(train,test)
comb_data$label=as.factor(comb_data$label)
```

Data Preprocessing
```{r}
# All pixels having 0 variance are removed
store=c(1)
for(i in 2:785){
  if (var(comb_data[,i])==0){store=append(store,i)}
}
comb_data=comb_data[,-store]
```
Dimensionality reduction by PCA (Principal Component Analysis)
```{r}
pca.train=comb_data[1:nrow(train),]
pca.test=comb_data[42001:nrow(comb_data),]
prin_comp=prcomp(pca.train)
std_dev = prin_comp$sdev
pr_var = std_dev^2
prop_varex = pr_var/sum(pr_var)
```
Visualizing Principal Components
```{r}
plot(prop_varex, xlab = "Principal Component",ylab = "Proportion of Variance Explained")
plot(cumsum(prop_varex), xlab = "Principal Component",ylab = "Cumulative Proportion of Variance Explained",type = "b")
```
Obtaining training and test data with variables as principal components
```{r}
train.data=data.frame(label=train$label,predict(prin_comp,pca.train))
test.data=data.frame(predict(prin_comp,pca.test))

rownames(test.data)=NULL
head(test.data)
train.data$label=as.factor(train.data$label)
```
Optimizing number of components to be selected
```{r}
#SVM = function(x){
#    obj=svm(label~.,data=train.data[,1:x])
#    return (predict(obj,newdata=train.data[1:100,2:x]))
#}
#no_of_components=c(10,25,50,100)
#tested=data.frame(labels=train.data$label[1:100])
#for (i in no_of_components){
#    mylabels=SVM(i)
#    tested=cbind(tested,mylabels)
#    paste(i,sum(diag(table(tested$labels,tested$mylabels))))
#}
```


Applying SVM alongwith tuning of hyperparameters on the chosen components(50) 
```{r}
svm_tune=tune(svm,label~.,data=train.data[sample(1:42000,1000),1:51],kernel="radial",ranges=list(cost=10^(-1:2),gamma=2^(-6:1)))
str(svm_tune)

# Optimal cost=10, gamma=0.0325
obj=svm(label~.,data=train.data[,1:51],cost=10,gamma=0.0325)       ## 50 principal components 
subm = data.frame(ImageId=1:nrow(test),Label=predict(obj,newdata=test.data[,1:50]))
write.csv(subm,'/kaggle/working/svmsub.csv',row.names=F)
```
